---
title: "p8105_hw6_bac2214"
author: "Brianna Carnagie"
date: "2023-11-28"
output: github_document
---
|
```{r, message=FALSE, echo=FALSE}
library(readr)
library(dplyr)
library(ggplot2)
library(modelr)
library(purrr)
library(tidyr)
library(broom)
```

## Question 1
In this first step, here is what I did:

 * A city_state variable was created.
 * A binary variable homicide_solved was added to indicate whether the homicide is solved.
 * Homicides from the cities Dallas, TX; Phoenix, AZ; Kansas City, MO were omitted. 
 * The analysis was limited to cases where victim_race is white or black.
 * victim_age was converted to a numeric format (and non-numeric values were coerced into NAs).
```{r setup, message=FALSE}
homicide_df = read_csv("data/homicide-data.csv") |> 
   mutate(
    city_state = paste(city, state, sep = ", "),
    homicide_solved = ifelse(grepl("closed", disposition, ignore.case = TRUE), 1, 0)
  ) |> 
  filter(!city_state %in% c("Dallas, TX", "Phoenix, AZ", "Kansas City, MO", "Tulsa, AL")) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  mutate(victim_age = as.numeric(victim_age))
```

In the next step, I attempt to make the log regression for Baltimore, MD and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing male victims to female victims keeping all other variables fixed.

```{r}
baltimore_df = filter(homicide_df, city_state == "Baltimore, MD")

baltimore_log_model = glm(homicide_solved ~ victim_age + victim_sex + victim_race, 
                       data = baltimore_df, family = binomial())

baltimore_log_model |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate),
    CI_lower = exp(estimate - 1.96 * std.error),
    CI_upper = exp(estimate + 1.96 * std.error)
  ) |> 
  select(term, OR, CI_lower, CI_upper) |> 
  knitr::kable(digits = 3)
```

**Key findings:**

* Intercept *(not interpretable)*: The odds of a homicide being solved when the victim is a 0-years old, black female is 1.92.
* For every one year increase in age, we expect the odds of their homicide being solved to decrease by 0.995
* The odds of a male victim having their homicide solved is 0.355 times the odds of a women having their homicide solved, holding all other factors constant.
* Being white significantly increases the odds of their homicide being solved, compared to being black.
* Victim age does not appear to be a significant predictor at the common 0.05 significance level (p>0.05).


Here I iterate the process to run a glm for all cities in my dataset:
```{r}

fit_glm_OR = function(df) {
  model = glm(homicide_solved ~ victim_age + victim_sex + victim_race, 
               data = df, 
               family = binomial())
  
  tidy_model = broom::tidy(model) |> 
    mutate(
      OR = exp(estimate),
      CI_lower = exp(estimate - 1.96 * std.error),
      CI_upper = exp(estimate + 1.96 * std.error)
    ) |> 
    filter(term == "victim_sexMale") |> 
    select(term, OR, CI_lower, CI_upper)
  return(tidy_model)
}

cities_OR_CI <- homicide_df %>%
  split(.$city_state) %>%
  map_dfr(fit_glm_OR, .id = "city_state")
```

Here's a plot:
```{r}
ggplot(cities_OR_CI, aes(x = reorder(city_state, OR), y = OR)) +
  geom_point() +
  geom_errorbar(aes(ymin = CI_lower, ymax = CI_upper), width = 0.2) +
  coord_flip() +  # Flips the axes for horizontal orientation
  labs(title = "Estimated Odds Ratios for Solving Homicides (Male vs Female Victims)",
       x = "City",
       y = "Estimated Odds Ratio")

```

Fresno has the highest odds of solving homicides comparing male victims to female victims, while New York has the lowest odds. Fresno has a really wide CI, indicating greater variability in the data.

## Question 2
```{r,message=FALSE}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())

suppressWarnings({
bootstrap_function <- function(data) {
  model <- lm(tmax ~ tmin + prcp, data = data)
  r_squared <- broom::glance(model)$r.squared
  log_beta_product <- log(broom::tidy(model)[2, "estimate"] * broom::tidy(model)[3, "estimate"])
  return(data.frame(r_squared = r_squared, log_beta_product = log_beta_product))
}

set.seed(123)  
bootstrap_estimates <- replicate(5000, sample_n(weather_df, size = n(), replace = TRUE), simplify = FALSE) %>%
  map_dfr(bootstrap_function) 
})
```

Here are some plots for Bootstrap Distribution of R-squared & Distribution of Bootstrap Estimates of log(beta1 * beta2):
```{r}
ggplot(bootstrap_estimates, aes(x = r_squared)) +
  geom_histogram( fill = "pink2", alpha = 0.7) +
  labs(title = "Bootstrap Distribution of R-squared", x = "R-squared", y = "Frequency")

ggplot(bootstrap_estimates, aes(x = estimate)) +
  geom_histogram(fill = "green3") +
  labs(title = "Distribution of Bootstrap Estimates of log(beta1 * beta2)",
       x = "log(beta1 * beta2) Estimate")

r_squared_ci <- quantile(bootstrap_estimates$r_squared, probs = c(0.025, 0.975))
log_beta_product_ci <- quantile(bootstrap_estimates$estimate, probs = c(0.025, 0.975), na.rm = TRUE)
```

* The Bootstrap Distribution of R-squared is slightly right-skewed while 
* The CI for r_squared is `r sprintf("%.2f", r_squared_ci[1])` to `r sprintf("%.2f", r_squared_ci[2])`, 
* The Bootstrap Distribution of log(beta1 * beta2) is skewed right significantly. 
* The CI for log(beta1 * beta2) is `r sprintf("%.2f", log_beta_product_ci[1])` to `r sprintf("%.2f", log_beta_product_ci[2])`.

## Question 3

I started this question off by importing the dataset and making factors for the categorical variables! 
```{r}
birthweight_df = read_csv("data/birthweight.csv") |> 
  mutate(
    babysex = factor(babysex, levels = 1:2, labels = c("Male", "Female")),
    frace = factor(frace, levels = c(1, 2, 3, 4, 8, 9), labels = c("White", "Black", "Asian", "Puerto Rican", "Other", "Unknown")),
    mrace = factor(mrace, levels = c(1, 2, 3, 4, 8), 
                   labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
    malform = factor(malform, levels = 0:1, labels = c("Absent", "Present"))
  ) |> 
  na.omit() 
```

This is my proposed regression model for birthweight.
```{r}

my_model <- lm(bwt ~ blength * gaweeks + ppbmi * smoken + momage * parity + wtgain, data = birthweight_df)

my_model_summary_stats= broom::tidy(my_model) 

```

* gaweeks: The duration of the pregnancy is a strong predictor of birth weight; full-term babies typically weigh more than pre-term babies.

* smoken: Smoking has been linked to lower birth weight, so including this variable can help account for its negative effect on the baby's weight.

* momage: Maternal age can be a factor in birth weight, with teen mothers and older mothers often having different risks for low birth weight.

* parity: The number of previous pregnancies can affect birth weight, as the uterus environment changes with each pregnancy.

* wtgain: Adequate weight gain during pregnancy is critical for fetal growth. Insufficient or excessive weight gain can both negatively affect birth weight.

* My interaction term accounts for the fact that the effect of gestational age on birth weight might differ for babies of smokers compared to non-smokers.

Below is my plot of residuals against fitted values
# Model diagnostics: 
```{r}
birthweight_data = birthweight_df |> 
  add_predictions(my_model, var = "pred") |> 
  add_residuals(my_model, var = "resid")

ggplot(birthweight_data, aes(x = pred, y = resid)) +
  geom_point(alpha = 0.5) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(title = "Residuals vs Fitted Values", x = "Fitted Values", y = "Residuals")
```

Here I compare my proposed model to two other models in terms of cross-validated prediction error:
```{r}
model1 <- lm(bwt ~ blength + gaweeks, data = birthweight_data)

model2 <- lm(bwt ~ bhead * blength * babysex, data = birthweight_data)

```




```


